---
title: "weather_type_prediction"
author: "Ting Chia Liu"
date: "2024-11-06"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
geometry: margin=1in
editor_options:
  markdown:
    wrap: 72
---

## Project Overview: Weather Type Classification

The goal of this project is to predict the type of weather based on a
range of 10 different features, including various environmental factors
such as temperature, humidity, wind speed, and more. These measurements
are critical for understanding weather patterns and can be used for
various applications in meteorology, agriculture, and other domains.

### Dataset Overview

The data used for this project is sourced from Kaggle
(<https://www.kaggle.com/datasets/nikhil7280/weather-type-classification/data>)
and contains measurements related to weather. It is a **multivariate
classification problem** where we aim to predict the weather type based
on several predictor variables. The dataset consists of continuous and
categorical variables.

Here's variables in the dataset:

1.  **Temperature (numeric)**: Represents the temperature in degrees
    Celsius.

2.  **Humidity (numeric)**: Represents the humidity percentage.

3.  **Wind Speed (numeric)**: Represents the wind speed in kilometers
    per hour (km/h).

4.  **Precipitation (%) (numeric)**: Represents the percentage of
    precipitation, indicating the likelihood of rain or snow.

5.  **Cloud Cover (categorical)**: Describes the amount of cloud cover
    in the sky such as "clear", "cloudy", "overcast", "partly cloudy"

6.  **Atmospheric Pressure (numeric)**: Represents the atmospheric
    pressure in hPa (hectopascals).

7.  **UV Index (numeric)**: Measures the intensity of ultraviolet
    radiation.

8.  **Season (categorical)**: Indicates the season during which the data
    was recorded.

9.  **Visibility (km) (numeric)**: Represents the visibility in
    kilometers (km).

10. **Location (categorical)**: Represents the type of location where
    the data was recorded, such as "inland", "mountain", "coastal".

11. **Weather Type (categorical, target variable)**: The target variable
    for classification, representing the type of weather at a given
    time. Categories include "Sunny", "Rainy", "Snowy", "Cloudy".

### Objective

The aim is to predict the **Weather Type** based on the 10 input
variables. This is a classification problem, and the most commonly used
evaluation metrics for classification include: - **Accuracy** -
**Confusion Matrix** - **Precision, Recall, F1-Score** -
**Cross-Validation Performance**

### Approach Overview

1.  **Data Preprocessing**:

    -   **Missing Data Handling**: Identify missing values.
    -   **Outlier Detection**: Handle outliers in numerical data (e.g.,
        temperature, humidity).
    -   **Feature Scaling**: Standardize or normalize numerical features
        like temperature, humidity, and wind speed to ensure that all
        features have a similar scale.

2.  **Exploratory Data Analysis (EDA)**:

    -   Visualize the distribution of each feature (e.g., histograms,
        box plots).
    -   Analyze correlations between numerical features.
    -   Investigate class distributions for the target variable (Weather
        Type).

3.  **Model Selection & Model Tuning**:

    -   Since this is a classification task, we used three algorithms:
        -   K-Nearest Neighbors (KNN): the value of `k` in KNN
        -   Decision Tree: the number of cp
        -   Random Forest: the number of mtry in a random forest
    -   Use cross-validation to assess model performance and select the
        best algorithm.
    -   Plot confusion matrix to visualize classification performance.
    -   Assess performance using metrics like accuracy, precision,
        recall, and F1-score.
    -   Use a separate test set to evaluate model performance.
    -   Comparison Model's Performance

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 8, 
  fig.height = 5, 
  out.width = '80%', 
  dpi = 300, 
  echo = TRUE
)
```

```{r Setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(randomForest)
library(class)
library(caret)
library(plyr)
library(dplyr)
library(psych)
library(pROC)
```

### 1. Data preparation

```{r}
weather <- read.csv("weather_classification_data.csv")
```

```{r}
# check data type
str(weather)
```

```{r}
# Changed the variable names (Precipitation and Visibility.km)
# and the character variables into factor. 
weather <- weather %>%
  dplyr::rename(
    Precipitation = `Precipitation....`,
    Visibility.km = `Visibility..km.`
  ) %>%
  mutate(
    Cloud.Cover = as.factor(Cloud.Cover),
    Season = as.factor(Season),
    Location = as.factor(Location),
    Weather.Type = as.factor(Weather.Type)
  )
```

```{r}
#check missing values
any(is.na(weather)) #no missing value
```

### 2.Exploratory Data Analysis (EDA)

#### 2.1 Description Dataset - summary function for each variable

```{r}
summary(weather)
```

#### 2.2 Target variable distirbution

There are four weather types in dataset: Cloudy, Rainy, Snowy and Sunny,
each with 3300 observations, indicating an even distribution across
weather categories.

```{r}
weather %>% 
  ggplot(aes(x = Weather.Type)) +
  geom_bar(aes(fill = Weather.Type))+
  geom_text(stat = "count", 
            aes(label = ..count..), 
            vjust = -0.5) +
  labs(title = "Distribution of Weather Type")+
  theme_minimal()

# For each weather type, there are 3300 obervations, 
# the data is scaled already, no need of data scaling.
```

#### 2.3 Quantitative variable distribution for different Weather Type

The distribution of humidity, precipitation, temperature, UV index, and
visibility vary significantly across the different weather types. We can
anticipate that these variables will greatly contribute for
classification of wheather type. On the other hand the pressure and wind
speed remain relatively consistent throughout the weather type, and will
probably not greatly influence the model for weather classification.
Humidity levels are obviously high during cloudy and rainy weather
types. Precipitation is much higher on rainy days and is nearly absent
on sunny days. Temperature, UV index, and visibility are highest on
sunny days. In contrast, pressure and wind speed do not shows
substantial variation across different weather types.

```{r}
# Distribution of the numerical variable based on 
# the different weather conditions
weather %>%
  pivot_longer(where(is.numeric)) %>%
  ggplot(aes(y = value, fill=Weather.Type)) +
  geom_boxplot(alpha = 0.8, color="Black") +
  facet_wrap(~name, scales = "free") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal()
```

#### 2.4 Qualitative variable distribution for different Weather Type

For the cloud cover variables, sunny weather is mostly associated with
clear cloud cover, while cloudy and rainy weather occurs mainly under
cloudy and overcast cloud cover. Rainy and sunny weather are equally
shown across coastal, inland, and mountain areas. While snowy weather is
mostly found in mountain and inland areas. Looking at the season
variables, snowy weather only occurs in winter, while sunny, rainy and
cloudy weather can be found in all seasons.

```{r}
weather %>%
  pivot_longer(cols = -c(where(is.numeric), Weather.Type)) %>%
  ggplot(aes(x = value, fill =Weather.Type)) +
  geom_bar(position = "dodge", 
           alpha = 0.9, 
           color = "black", 
           size = 0.3) +
  facet_wrap(~name, scales = "free") +
  scale_fill_viridis_d() + 
  theme_minimal(base_size = 14) +
  labs(
    title = "Distribution of Categorical Variables by Weather Conditions") +
  theme(
    legend.position = "top",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### 3. Model Selection & Model Tuning

```{r}
set.seed(123) #For reproducibility
# create a train_index to help divide the train and test dataset
train_index <- sample(1:nrow(weather), 0.8*nrow(weather))

train_weather <- weather[train_index,]
test_weather <- weather[-train_index,]
```

#### 3.1 KNN (tuning K value)

```{r, fig.width=5, fig.height=4, fig.align="center"}
set.seed(123) #For reproducibility
#KNN with Z-score scaling
train_control <- trainControl(method = "cv", number = 10)
#define K range from 1 to 20
tune_grid <- expand.grid(k = 1:20)

knn_fit <- train(
  Weather.Type ~ .,
  data = train_weather,
  method = "knn",
  trControl = train_control,
  tuneGrid = tune_grid,
  preProcess = c("center", "scale")
)

plot(knn_fit)

print(knn_fit)

knn_fit$bestTune #Best K=3
```

After tuning the value of k from 1 to 20, the plot indicates that the
highest accuracy is achieved when k=3, suggesting that this value of k
provides the best model performance.

#### Prediction with KNN

```{r}
predictions_knn <- predict(knn_fit, newdata = test_weather)
cm <- confusionMatrix(predictions_knn, test_weather$Weather.Type)
cm

# Plot confusion matrix as a heatmap
cm_df <- as.data.frame(cm$table)
ggplot(cm_df, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "black") +
  scale_fill_gradient(low = "pink", high = "red") +
  labs(title = "Confusion Matrix Heatmap",
       x = "Predicted Label",
       y = "Observed Label")

# Create a data frame to compare actual and predicted classes
comparison_df <- data.frame(Actual = test_weather$Weather.Type, 
                            Predicted = predictions_knn)

# Create a bar plot of Actual vs Predicted classes
ggplot(comparison_df, aes(x = Actual, fill = Predicted)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count", aes(label = ..count..), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5) +
  labs(title = "Actual vs Predicted Weather Types",
       x = "Actual Weather Type",
       y = "Count") +
  theme_minimal()
# The bigger bars are the correctly predicted classes and 
# the small bars are the misclassified predictions.
```

The number of samples in each class of the testing set is 660. Both the
'Cloudy' and 'Snowy' classes show better prediction performance.

#### 3.2 Classification Tree (tuning cp value)

```{r Decision Tree with cross validation, fig.width=5, fig.height=4, fig.align="center"}
set.seed(123)
# Find optimal complexity and tree size
mse_prune <- c()
i <- 1
tree_model <- rpart(Weather.Type ~ ., 
                    train_weather, 
                    method="class", 
                    cp=0.01, 
                    control=rpart.control(xval=10))
for (cp in seq(0, 0.1, by=0.01)){
  model <- prune(tree_model, cp=cp) 
  predicted <- predict(model, test_weather, type="class")
  mse_prune[i] <- mean(test_weather$Weather != predicted)
  i <- i+1 # don't forget this one
}
tibble(mse=mse_prune, cp=seq(0, 0.1, by=0.01)) %>% 
  ggplot(aes(x=cp, y=mse)) + 
  geom_point()+ 
  geom_line() + 
  theme_minimal() # best is 0.01


rpart.plot(tree_model)

#could also be with rpart control
plotcp(tree_model) # see relationship between tree size, CP and relative error
printcp(tree_model) # see relationship between tree size, CP and relative error
```

When CP = 0.01 (with 5 splits, as shown in the tree plot), the model
achieves the lowest cross-validated error, but it results in a more
complex model.

#### Prediction on Classification Tree

```{r}
predictions_tree <- predict(tree_model, newdata = test_weather, type="class")

confusionMatrix(predictions_tree, test_weather$Weather.Type) #accuracy 88.18%


# Plot feature importance using ggplot2
importance_data <- data.frame(
  Feature = names(tree_model$variable.importance),
  Importance = tree_model$variable.importance
)

ggplot(importance_data, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Feature Importance from Decision Tree Model", 
       x = "Feature", 
       y = "Importance") +
  theme_minimal() +
  coord_flip()  # Flip coordinates to make the bars horizontal

# Create a data frame with true and predicted labels
results_df <- tibble(True = test_weather$Weather, 
                     Predicted = predictions_tree)

# Faceted bar plot of True vs Predicted
ggplot(results_df, aes(x = Predicted, fill = True)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count", aes(label = ..count..), 
            position = position_dodge(width = 0.8), 
            vjust = 0.5) + 
  facet_wrap(~ True) +
  labs(title = "True vs. Predicted Weather Types", 
       x = "Predicted Weather Type", y = "Count") +
  theme_minimal()

# Visualize the splitting pattern of a decision tree model 
# based on two features 
ggplot(test_weather, 
       aes(x = Temperature, y = Humidity, 
           color = predictions_tree)) +
  geom_point(alpha = 0.6) +
  labs(title = "Predicted Weather Types by Temperature and Humidity", 
       x = "Temperature", y = "Humidity", color = "Predicted") +
  theme_minimal()
```

From the optimized Decision Tree, we achieved an accuracy of 88.18%. The
top three most important variables are: Temperature, Precipitation, and
Pressure. Notably, the 'Snowy' and 'Rainy' classes exhibit better
prediction performance compared to other weather types. The scatter plot
shows the distribution of data points based on Temperature and Humidity.
We observe that data points for each class are more condensed along the
Temperature axis (x-axis), indicating a clearer separation, whereas they
are more dispersed along the Humidity axis (y-axis), reflecting greater
variability in humidity across different weather types.

#### 3.3 Random Forest (tuning mtry)

```{r Random Forest with CV}
set.seed(123)
# Set up cross-validation with 5 folds
control <- trainControl(method = "cv", number = 5)

tune_grid <- expand.grid(mtry = 1:10)
# Perform cross-validation using the caret package
rf_cv <- train(Weather.Type ~ ., 
               data = train_weather, 
               method = "rf",
               trControl = control, 
               tuneGrid = tune_grid, 
               ntree = 500)

# Print the results
print(rf_cv)

# Extract the cross-validation results
cv_results <- rf_cv$results

# Plot the cross-validation error
ggplot(cv_results, aes(x = mtry, y = 1 - Accuracy)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Cross-Validation Error vs. mtry",
       x = "mtry",
       y = "CV Error (1 - Accuracy)") +
  scale_x_continuous(breaks = 1:10) 
#We got the highest acc when mtry=5 during tuning.


# Prediction on the test set
# Loop through mtry from 1 to 10 
# just for validation: Does mtry=5 have the higest accuracy on test set
accuracies <- numeric(10)
mtry_values <- 1:10

for (i in mtry_values) {
  set.seed(123)
  rf_mod <- randomForest(Weather.Type ~ ., 
                         data = train_weather, 
                         mtry = i, ntree = 500)
  predictions_rf <- predict(rf_mod, newdata = test_weather) 
  cm <- confusionMatrix(predictions_rf, test_weather$Weather.Type)
  accuracies[i] <- cm$overall['Accuracy']
  cat("mtry =", i, "Accuracy =", accuracies[i], "\n")
}

# Plot the accuracy of each mtry value
accuracy_df <- tibble(mtry = mtry_values, accuracy = accuracies)
ggplot(accuracy_df, aes(x = mtry, y = accuracy)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(
    title = "Accuracy vs mtry for Random Forest Model",
    x = "mtry",
    y = "Accuracy"
  ) +
  scale_x_continuous(breaks = 1:10)
```

#### Prediction with Random Forest mtry=5

```{r}
set.seed(123)
# Random forest with mtry=5
rf_mod_5 <- randomForest(Weather.Type ~ ., 
                         data = train_weather, 
                         mtry=5, ntree = 500)

#prediction on mtry5 model
predictions_rf_5 <- predict(rf_mod_5, newdata = test_weather)
head(predictions_rf_5)
confusionMatrix(predictions_rf_5, test_weather$Weather.Type)
# accuracy 91.67%

# Visualize the confusion matrix
cm_table <- as.table(cm$table)
cm_data <- as.data.frame(cm_table)
colnames(cm_data) <- c("Actual", "Predicted", "Freq")

ggplot(data = cm_data, aes(x = Actual, y = Predicted, fill = Freq)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  geom_text(aes(label = Freq), vjust = 1) +
  theme_minimal() +
  labs(
    title = "Confusion Matrix for Random Forest Model (mtry=5)",
    x = "Actual",
    y = "Predicted"
  )

#visualize variable importance
var_imp <- importance(rf_mod_5)

tibble(
  importance = c(var_imp), 
  variable = rownames(var_imp)
 )%>% 
  ggplot(aes(x = importance, y = reorder(variable,importance), 
             fill = variable)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d() +
  theme_minimal() 
```

The random forest model with mtry = 5 (best test mtry) achieved an
accuracy of 91.67%. The confusion matrix reveals that the classes
'Snowy' and 'Cloudy' have better prediction performance compared to
other classes. The top three most important variables contributing to
the model are: Temperature, Precipitation, and UV Index which different
than the ones for the classification trees for which: Temperature,
Precipitation, and Pressure were the top 3 most important variables.

#### Feature Selection: 
We selected the five most important features based on the random forest model with mtry =5 to train, tune, and evaluate its performance.

```{r}
# Set seed for reproducibility
set.seed(123)

# Define the range of mtry values to test (from 1 to 5)
mtry_values <- 1:5

# Initialize a variable to store the OOB error for each mtry
oob_errors <- numeric(length(mtry_values))

# Loop over each mtry value and fit a model, recording the OOB error
for (i in mtry_values) {
  model <- randomForest(
    Weather.Type ~ Temperature + Precipitation + 
      Visibility.km + UV.Index + Atmospheric.Pressure,
    data = train_weather,
    mtry = i,
    ntree = 500
  )
  # Store the OOB error for the current model
  oob_errors[i] <- model$err.rate[nrow(model$err.rate), "OOB"]
}

# Find the best mtry (with the lowest OOB error)
best_mtry <- mtry_values[which.min(oob_errors)]
best_mtry

# Print the best mtry value and its corresponding OOB error
cat("Best mtry:", best_mtry, "\n")
cat("OOB Error at best mtry:", min(oob_errors), "\n")

# Plot the OOB error for each mtry value
plot(mtry_values, oob_errors, type = "b", pch = 19, col = "blue",
     xlab = "mtry", ylab = "OOB Error Rate",
     main = "OOB Error Rate for Different mtry Values")
```

From the OBB plot, we chose mtry=2 which has shown to have the lowest
error rate for the top5 variable model.

```{r}
# Random forest with mtry=2 with the top 5 variable 
# based on the variable importance plot
set.seed(123)
rf_mod_top_5<-  randomForest( 
  Weather.Type ~ Temperature + Precipitation + 
    Visibility.km + UV.Index + Atmospheric.Pressure, 
  data = train_weather, 
  mtry = 2, 
  ntree = 500
)
#prediction on top 5 variables, mrty=2 model
predictions_rf_top_5 <- predict(rf_mod_top_5, newdata = test_weather)
head(predictions_rf_top_5)
confusionMatrix(predictions_rf_top_5, test_weather$Weather.Type)
# accuracy 0.89

conf_mat <- confusionMatrix(predictions_rf_top_5, test_weather$Weather.Type)

# Convert the confusion matrix to a data frame for ggplot
conf_mat_df <- as.data.frame(conf_mat$table)

# Plot the confusion matrix
ggplot(data = conf_mat_df, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "white", size = 5) +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  labs(title = "Confusion Matrix", x = "Predicted", y = "Actual") +
  theme_minimal()
```

The Random Forest model using only the top 5 variables and setting mtry
= 2 achieved an accuracy of 89.66%. While this is a good result, it is
lower than the model that included all the variables. This suggests
that, while the top 5 variables may be the most informative, the
additional, less important variables might still contribute to capturing
some underlying patterns in the data, improving the model’s overall
performance.

#### 3.4. Comparison Model's Performance

```{r Comparison of the performance, warning=FALSE, fig.width=5, fig.height=4, fig.align="center"}
#form correction
list_prediction <- list(predictions_knn, 
     predictions_tree,
     predictions_rf_5, 
     predictions_rf_top_5)

accuracy_list <- c()
for (i in (1:4)){
  cm<- confusionMatrix(list_prediction[[i]], test_weather$Weather)
  accuracy_list[i] <- cm$overall["Accuracy"]
}

tibble(Model=c("KNN", "Decision Tree", 
               "Random Forest mtry5",
               "Random Forest Top 5 Variables"), 
       Accuracy = accuracy_list) %>% 
  ggplot(aes(x=Model, y= Accuracy,  group = 1)) + 
  geom_point() + 
  geom_line() + 
  theme_minimal() + 
  geom_label(aes(x = "Random Forest mtry5", y = 0.95, label = "Best Model")) +  
  coord_cartesian(ylim = c(0.80, 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Overall, we evaluated several models: K-Nearest Neighbors (KNN),
Decision Tree, Random Forest with mtry = 5 and Random Forest using the
top 5 variables. The Random Forest model with mtry = 5 achieved the
highest accuracy.

Discussion of Model Performance Across Each Class: The following ROC
plots displays the performance of the four models, evaluating their
accuracy for each of the Weather class: sunny, cloudy, sunny and snowy,
```{r include=FALSE}
# Prediction if a day will be Sunny
set.seed(123)
# Get predicted probabilities for KNN model
knn_probs <- predict(knn_fit, newdata = test_weather, type = "prob")
# Get predicted probabilities for Decision Tree model
tree_probs <- predict(tree_model, newdata = test_weather, type = "prob")
# Train Random Forest model with mtry = 5
rf_mod_5 <- randomForest(y = train_weather$Weather.Type, 
                         x = train_weather[ ,-11], 
                         mtry = 4, 
                         ntree = 500)

# Get predicted probabilities for Random Forest with mtry = 5
rf_probs_5 <- predict(rf_mod_5, newdata = test_weather, type = "prob")

# Train Random Forest model with the top 5 features and mtry=2
rf_mod_top_5 <- randomForest(y = train_weather$Weather.Type, 
                         x = train_weather[ ,-11], 
                         mtry = 2, 
                         ntree = 500)
# Get predicted probabilities for Random Forest with mtry = 2 , top 5 variables
rf_probs_top_5 <- predict(rf_mod_top_5, newdata = test_weather, type = "prob")

# Generate ROC curves and calculate AUC for the different classes
# Assuming 'Sunny' is the class of interest
knn_roc <- roc(test_weather$Weather.Type, knn_probs[, "Sunny"])
tree_roc <- roc(test_weather$Weather.Type, tree_probs[, "Sunny"])
rf_roc_5 <- roc(test_weather$Weather.Type, rf_probs_5[, "Sunny"])
rf_roc_top_5<- roc(test_weather$Weather.Type, rf_probs_top_5[, "Sunny"])
```

```{r}
# Calculate AUC for each model
knn_auc <- auc(knn_roc)
tree_auc <- auc(tree_roc)
rf_auc_5 <- auc(rf_roc_5)
rf_roc_top_5 <- auc(rf_roc_top_5)

# Print AUCs
cat("KNN AUC:", knn_auc, "\n")
cat("Decision Tree AUC:", tree_auc, "\n")
cat("Random Forest (mtry=5) AUC:", rf_auc_5, "\n")
cat("Random Forest (mtry=2) Top 5 Variables AUC:", rf_roc_top_5, "\n")
# Plot the ROC curve for KNN as the initial plot
plot(knn_roc, col = "blue", main = "ROC Curve Comparison - Sunny", lwd = 2)

# Add other ROC curves to the same plot
lines(tree_roc, col = "green", lwd = 2)
lines(rf_roc_5, col = "red", lwd = 2)
lines(rf_roc_top_5, col = "grey", lwd = 2)

# Add legend
legend("bottomright", 
       legend = c("KNN", "Classification Tree", 
                  "RF (mtry=5)",  "RF Top 5 Variables"),
       col = c("blue", "green", "red", "grey"), 
       lwd = 2)
```
```{r include=FALSE}
# Generate ROC curves and calculate AUC for each class (Cloudy, Snowy, Rainy)

# ROC for Cloudy class
knn_roc_cloudy <- roc(test_weather$Weather.Type, knn_probs[, "Cloudy"])
tree_roc_cloudy <- roc(test_weather$Weather.Type, tree_probs[, "Cloudy"])
rf_roc_5_cloudy <- roc(test_weather$Weather.Type, rf_probs_5[, "Cloudy"])
rf_roc_top_5_cloudy <- roc(test_weather$Weather.Type, rf_probs_top_5[, "Cloudy"])

# ROC for Snowy class
knn_roc_snowy <- roc(test_weather$Weather.Type, knn_probs[, "Snowy"])
tree_roc_snowy <- roc(test_weather$Weather.Type, tree_probs[, "Snowy"])
rf_roc_5_snowy <- roc(test_weather$Weather.Type, rf_probs_5[, "Snowy"])
rf_roc_top_5_snowy <- roc(test_weather$Weather.Type, rf_probs_top_5[, "Snowy"])

# ROC for Rainy class
knn_roc_rainy <- roc(test_weather$Weather.Type, knn_probs[, "Rainy"])
tree_roc_rainy <- roc(test_weather$Weather.Type, tree_probs[, "Rainy"])
rf_roc_5_rainy <- roc(test_weather$Weather.Type, rf_probs_5[, "Rainy"])
rf_roc_top_5_rainy <- roc(test_weather$Weather.Type, rf_probs_top_5[, "Rainy"])
```

```{r}
# Cloudy ROC plot
plot(knn_roc_cloudy, col = "blue", 
     main = "ROC Curve Comparison - Cloudy", lwd = 2)
lines(tree_roc_cloudy, col = "green", lwd = 2)
lines(rf_roc_5_cloudy, col = "red", lwd = 2)
lines(rf_roc_top_5_cloudy, col = "grey", lwd = 2)
legend("bottomright", 
       legend = c("KNN", "Classification Tree", 
                  "RF (mtry=5)", "RF Top 5 Variables"),
       col = c("blue", "green", "red", "grey"), 
       lwd = 2)

# Snowy ROC plot
plot(knn_roc_snowy, col = "blue", 
     main = "ROC Curve Comparison - Snowy", lwd = 2)
lines(tree_roc_snowy, col = "green", lwd = 2)
lines(rf_roc_5_snowy, col = "red", lwd = 2)
lines(rf_roc_top_5_snowy, col = "grey", lwd = 2)
legend("bottomright", 
       legend = c("KNN", "Classification Tree", 
                  "RF (mtry=5)", "RF Top 5 Variables"),
       col = c("blue", "green", "red", "grey"), 
       lwd = 2)

# Rainy ROC plot
plot(knn_roc_rainy, col = "blue", 
     main = "ROC Curve Comparison - Rainy", lwd = 2)
lines(tree_roc_rainy, col = "green", lwd = 2)
lines(rf_roc_5_rainy, col = "red", lwd = 2)
lines(rf_roc_top_5_rainy, col = "grey", lwd = 2)
legend("bottomright", 
       legend = c("KNN", "Classification Tree", 
                  "RF (mtry=5)", "RF Top 5 Variables"),
       col = c("blue", "green", "red",  "grey"), 
       lwd = 2)
```

Conclusion: The Classification Tree model achieves the best AUC/ROC for
Sunny and Snowy days compared to the other models, while the three other
models (KNN and the two random Forests) struggle to categorize for these
two classes. However, for the remaining two classes, all models provide
strong predictive performance. The Classification tree has more
consistent predictive performance than the three other models based on
the ROC.
